---
title: "p8105_hw2_am4656"
author: "Aaron Mittel"
date: "2022-10-04"
output: github_document
---

```{r setup, echo=FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

_Imported "NYC_Transit_and_Subway_Entrance_and_Exit_Data.csv" from a relative file path to my working directory, cleaned column names, and then selected for only "line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance" by selecting for the range of columns incorporating those variables and removing unwanted variables. Then, converted the entry variable from character (YES vs NO) to a logical variable using "ifelse"._

```{r data input, echo = FALSE, message = FALSE}
transit_data = read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>%
  select(line:ada, -exit_only, -staffing, -staff_hours) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

Following these procedures, this dataset is now organized into **19 columns and 1868 rows**. The dataset contains variables describing individual tracks/lines of the New York city subway transit system by the associated street (e.g. 4th Avenue) that the track/line runs along, stations along each track/line, individual trains/routes (e.g. A, B, C, or D) that run on each track/line, geographic positions of each station, and basic features of each station (whether a vending machine is present, stairs or elevator used for access, whether the station is ADA compliant).

The data is tidier than it once was. However, it does have some redundancy and can be improved. Route # should be a variable and route should also be a variable (done below).

_We can select for unique combinations of station name, line, and ADA accesability using the "distinct" function"._

```{r, echo = FALSE, message = FALSE}
transit_data_distinct_stations = transit_data %>%
  select(line, station_name) %>%
  distinct
```

* There are **`r nrow(transit_data_distinct_stations)`** distinct stations.

```{r, echo = FALSE, message = FALSE}
transit_data_distinct_ada = transit_data %>%
  filter(ada == TRUE) %>%
  distinct(line, station_name)
```

* There are **`r nrow(transit_data_distinct_ada)`** distinct stations that are ADA compliant.

_We can compute the proportion of station entrances without vending that allow entrance by first excluding station entrances that do not allow vending. Then, we can take the mean of the entry variable (which will produce the desired proportion because R will coerce logical to numeric in cases like this)._

```{r message=FALSE, include=FALSE}
transit_data %>%
  filter(vending == "NO") %>%
  pull(entry) %>% 
  mean()
```

* The proportion of station entrances without vending that allow entrance is thus **`r (transit_data %>% filter(vending == "NO") %>% pull(entry) %>% mean)`**.

_We can identify distinct stations that serve the A train and are ADA compliant by tidying the data to convert `route` from wide to long format and filtering to focus on the A train._

```{r, echo = FALSE, message = FALSE}
transit_data %>% 
pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Thus, there are only **17** stations that serve the A train that are ADA compliant.


## Problem 2

_Reading Mr. Trash Wheel into the R Project, the self-named first sheet in the Excel file. We must skip the first row (which is just an image in the original Excel file) and the last row (which appears to be summed column values) to omit rows that do not include dumpster-specific data. Variable names are largely reasonable, though we must add a variable to distinguish this data from the additional Professor Trash Wheel data that will be added in the next step. We convert the `sports_balls` variable to an integer and also change the `year` variable to numeric type, instead of character, so that it may be joined with Professor Trash Wheel._

```{r mr trash wheel data input, echo = FALSE, message = FALSE}
mr_trash_wheel = read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = cell_rows(2:549)) %>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered) %>%
  mutate(wheel = "mister") %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(year = as.numeric(year))
```

_Reading Professor Trash Wheel into the R Project in a comparable manner as above._

```{r professor trash wheel data input, echo = FALSE, message = FALSE}
dr_trash_wheel = read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = cell_rows(2:96)) %>%
  janitor::clean_names() %>% 
  select(dumpster:homes_powered) %>%
  mutate(wheel = "doctor")
```

_Stacking Mr. Trash Wheel and Professor Trash Wheel and relocating the `wheel` and `dumpster` variables to the left side of the dataset for easy viewing._

```{r stacking mr and dr trash wheel, echo = FALSE, message = FALSE}
trash_wheels = 
  bind_rows(mr_trash_wheel, dr_trash_wheel) %>%
  relocate(wheel,dumpster)

trash_wheels
```

There are **`r ncol(trash_wheels)` columns** and **`r nrow(trash_wheels)` rows** in this combined dataset.

Key variables of interest include the specific trash wheel, either `mister` or `doctor`, `weight_tons`, and quantities of commonly collected items, including `cigarette_butts`, `chip_bags`, and other similar items. It is worth noting that `sports_balls` are collected (or recorded) by Mr. Trash Wheel but not by Professor Trash Wheel.

* The total weight of trash collected by Professor Trash Wheel was **`r sum(trash_wheels %>% filter(wheel == "doctor") %>% pull(weight_tons))` tons**.

* The total number of sports balls collected by Mr. Trash Wheel in 2020 was **`r trash_wheels %>% filter(wheel == "mister", year == "2020") %>% pull(sports_balls) %>% sum`**.


# Problem 3
