---
title: "p8105_hw2_am4656"
author: "Aaron Mittel"
date: "2022-10-04"
output: github_document
---

```{r setup, echo=FALSE, message = FALSE}
library(tidyverse)
library(dplyr)
```

# Problem 1

_Imported "NYC_Transit_and_Subway_Entrance_and_Exit_Data.csv" from a relative file path to my working directory, cleaned column names, and then selected for only "line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance" by selecting for the range of columns incorporating those variables and removing unwanted variables. Then, converted the entry variable from character (YES vs NO) to a logical variable using "ifelse".

```{r data input, echo = FALSE, message = FALSE}
transit_data =
  read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line:ada, -exit_only, -staffing, -staff_hours) %>%
```

Following these procedures, this dataset is now organized into **19 columns and 1868 rows**. The dataset contains variables describing individual tracks/lines of the New York city subway transit system by the associated street (e.g. 4th Avenue) that the track/line runs along, stations along each track/line, individual trains/routes (e.g. A, B, C, or D) that run on each track/line, geographic positions of each station, and basic features of each station (whether a vending machine is present, stairs or elevator used for access, whether the station is ADA compliant).

The data is tidier than it once was. However, it does have some redundancy and can be improved.



